---
title: "STAT 3280"
author: "Jme6bk"
date: "8/31/2022"
output: html_document
---
```{r}
setwd("/Users/jillianeberhart/Desktop/Github/STAT-3280-Data-Visualization/STAT 3280 - 08:31.Rmd")
my_theme <- theme_bw() + theme(axis.text = element_text(size = 10), axis.title = element_text(size = 12))
library(dplyr)
library(ggplot2)
library(tidyverse)
library(scales)
```


##*Pie Charts*  
- Area proportional to quantity  
- categorical or grouped variables  
- Comparison between categories  
- Provides no information on quantity, just proportion  
- Difficult to distinguish small differences  
- useful as long as the number of categories isn't too large

```{r}
eye_data <- class %>% count(eyecolor) %>% mutate(frac = n / sum(n))
class$eyecolor[class$eyecolor == "dark brown"] <- "brown"
class$eyecolor[class$eyecolor == "eye color"] <- NA

eye_data <- drop_na(class) %>% count(eyecolor) %>%
  mutate(frac = n / sum(n))
```

*code*
- geom_bar : single variable, but we specify too --> stat = "identity"
- "bar" = "one slice of circle"
- coord.polar --> coordinate adjustment make it a pie chart
  + direction : counter or clockwise
  + start : where to start (always 0)
```{r}
plot2.1 <- ggplot(eye_data, aes(x = "", y = frac, fill = eyecolor)) +
  geom_bar(width= 1, stat = "identity", color = "black") +
  coord_polar("y", start = 0, direction = -1)
```

###This could look better - lets fix
- add theme_void : cleans up Axis
- re order quantities (in descending order)

```{r}

plot2.1 <- ggplot(eye_data, aes(x = "", y = frac, fill = reorder(eyecolor,l -frac))) +
  geom_bar(width= 1, stat = "identity", color = "black") +
  coord_polar("y", start = 0, direction = -1) + theme_void()

#this is going in counter clockwise direction
```

- fix colors (scale command)
- make it softer (alpha command - semi transparent)
  + to make one transparent (alpha("darkgreen", 0.5))
```{r}
plot2.2 <- plot2.1 + scale_fill_manual("Eye Color", values = alpha(c("green" = "darkgreen", "black" = "black", "brown" = "darkorange4", "blue" = "steelblue3", "hazel" = "darkolivegreen4"), 0.5))
```

- annotate our plot with proportions(specify labels)
  + if your going to annoate a transformed coordinate system to be able to tell R where to print specific labels

```{r}
eye_data <- eye_data %>% arrange(desc(n)) %>% mutate(prop = round(100 * frac, 0), labpos = 100 - (cumsum(prop) - 0.5 * prop))

```

###FINAL with not the best labels ###
```{r}
plot2.3 <- ggplot(eye_data, aes(x = "", y = prop, fill = reorder(eyecolor, -prop))) +
  geom_bar(width = 1, stat = "identity", color = "black") +
  geom_text(aes(y = labpos, label = paste(prop, "%", sep = ""))) +
  coord_polar("y", start = 0, direction = 1) + 
  scale_fill_manual("Eye Color", values = alpha(c("green" = "darkgreen", "black" = "black", "brown" = "darkorange4", "blue" = "steelblue3", "hazel" = "darkolivegreen4"), 0.5)) +
  theme_void()
plot2.3
  
```

- Annotate with title
```{r}
plot2.5 <- ggplot(eye_data, aes(x = "", y = prop, fill = reorder(eyecolor, -prop))) +
  geom_bar(width = 1, stat = "identity", color = "black") +
  geom_text(aes(y = labpos, label = paste(str_to_title(eyecolor), "\n", prop, "%", sep = ""))) +
  coord_polar("y", start = 0, direction = 1) + 
  scale_fill_manual("Eye Color", values = alpha(c("green" = "darkgreen", "black" = "black", "brown" = "darkorange4", "blue" = "steelblue3", "hazel" = "darkolivegreen4"), 0.5)) +
  theme_void() +
  theme(legend.position = "none") + 
  labs(title = "Eye Color in STAT 3280")
plot2.5
```

- Annotate with count
```{r}
plot2.6 <- ggplot(eye_data, aes(x = "", y = prop, fill = reorder(eyecolor, -prop))) +
  geom_bar(width = 1, stat = "identity", color = "black") +
  geom_text(aes(y = labpos, label = paste(str_to_title(eyecolor), "\n", prop, "%", "\n n = ", n, sep = ""))) +
  coord_polar("y", start = 0, direction = 1) + 
  scale_fill_manual("Eye Color", values = alpha(c("green" = "darkgreen", "black" = "black", "brown" = "darkorange4", "blue" = "steelblue3", "hazel" = "darkolivegreen4"), 0.5)) +
  theme_void() +
  theme(legend.position = "none") + 
  labs(title = "Eye Color in STAT 3280")
plot2.6
```

- declutter
```{r}
plot2.6 <- plot2.6 + labs(title = paste("Eye Color in STAT 3280, n = ", sum(eye_data$n), " students", sep = ""))
plot2.6
```


##*Dot Plots*  
- Display uni-variate categorical or quantitative data:
- horizontal or vertical
- number of dots equal (proportional to ) quantity
- categorical, grouped, or binned quantitative variables
- useful as long as number of dots is not too large
- usually can be replaced by another type of plot where the message is more clear (histogram/bar plot)
- difficult to work with on coordinate system in ggplot

```{r}
plot3.1 <- ggplot(class) +
  geom_dotplot(aes(x = music)) +
  labs(x = "Favorite Music Genre", y = "Proportion") +
  my_theme
plot3.1
```


- change dot size
```{r}
plot3.2 <- ggplot(class) +
  geom_dotplot(aes(x = music), dotsize = 2) +
  labs(x = "Favorite Music Genre", y = "Proportion") +
  my_theme
plot3.2
```

- adjust scale
  + coord_fix : modify with ratio
  + scale_y_continuous : tell ggplot what to print out on the side of coordinate system
  + 
```{r}
max_value <- max(count(class, music)["n"])
plot3.2 <- ggplot(class) +
  geom_dotplot(aes(x = music), dotsize = 1, fill = "lightblue", color = "black") +
  coord_fixed(ratio = 8.4) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 1 / max_value), labels = seq(0, max_value)) +
  labs(x = "Favorite Music Genre", y = "Count") +
  my_theme +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1))
plot3.2
```

- Quantitative Dot Sizes
```{r}
plot3.3 <- ggplot(class) +
  geom_dotplot(aes(x = hours), method = "histodot") +
  labs(x = "Hours Enrolled", y = "Proportion") +
  my_theme
plot3.3
```

```{r}
max_value <- max(count(class, hours)["n"])

plot3.4 <- ggplot(class) +
  geom_dotplot(aes(x = hours), method = "histodot", binwidth = 1, dotsize = 1) +
  coord_fixed(ratio = max_value) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 1 / max_value), labels = seq(0, max_value)) +
  labs(x = "Hours Enrolled", y = "Count") +
  my_theme
plot3.4
```

- Change Stackdir
```{r}
plot3.4 <- ggplot(class) +
  geom_dotplot(aes(x = hours), method = "histodot", binwidth = 1, dotsize = 1, stackdir = "down") +
  coord_fixed(ratio = max_value) +
  scale_y_continuous(limits = c(-1, 0), breaks = seq(0, -1, -1 / max_value), labels = seq(0, max_value)) +
  labs(x = "Hours Enrolled", y = "Count") +
  my_theme
plot3.4
```

- Change stacked ratio

```{r}
plot3.5 <- ggplot(class) +
  geom_dotplot(aes(x = hours), method = "histodot", binwidth = 1, dotsize = 1, stackratio = 0.5, fill = "lightblue") +
  coord_fixed(ratio = 0.5 * max_value) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 1 / max_value), labels = seq(0, max_value)) +
  labs(x = "Hours Enrolled", y = "Count") +
  my_theme
plot3.5

plot3.6 <- ggplot(class) +
  geom_dotplot(aes(x = hours), fill = "lightblue", color = "black", method = "histodot", binwidth = 1, dotsize = 1) +
  coord_fixed(ratio = max_value) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 1 / max_value), labels = seq(0, max_value)) +
  labs(x = "Hours Enrolled", y = "Count") +
  my_theme
plot3.6
```


- 